\begin{abstract}
The BitTorrent protocol is a popular peer-to-peer algorithm that is able
to transfer large files efficiently over the internet. A core component
of the algorithm is the local rarest first piece selection algorithm,
which selects pieces to download based on rarity with a node's peer
set to maximize the number of copies of each piece in a BitTorrent swarm. 
In this paper, we study how good the local knowledge within the peer set is compared to
complete global knowledge of piece rarity. The hypothesis is that piece
selection with perfect global knowledge can either improve torrent
performance and/or extend the life of unhealthy swarms.  We implemented a
BitTorrent simulator that allows nodes to access complete global knowledge
at any time and ran experiments. We find that for healthy swarms with seeders
that complete global knowledge does not improve the average download
times for nodes. However, for unhealthy swarms that have no seeders
but still have a complete copy of the file distributed in the swarm,
we find that complete global knowledge can help prevent the swarm from
losing pieces, which results in nodes being unable to complete their downloads. 
Next, we design and analyze a gossip protocol that can approximate complete
global knowledge efficiently. We find that gossip can help swarms with no seeders
almost as much as complete global knowledge. We also
describe an alternative gossip protocol to help seeders determine if it is
safe to leave a swarm without causing pieces to be lost. 

\end{abstract}
