\section{Future Work}

There are three main areas of future work for this project: adding realism 
to the simulation, deploying the gossip protocol in a BitTorrent client,
and exploring further uses for a low-overhead gossip protocol. 

Currently, our simulation implements all of the core features of
BitTorrent necessary to study the utility of global knowledge and
gossip. The only feature that might affect performance that we did not
implement was super-seeding, which we felt would not change our overall
results. We would like to investigate running swarm simulations with
more realistic features though such as a realistic churn rate and a more
diverse set of downloaders using a real distribution of common BitTorrent
download rates. In addition, while the current ``unhealthy'' swarm test is
effective, a more realistic test might be to just have the swarm's only
seed suddenly drop after some specified time. This seems even more likely to
occur in real work situations. A swarm with our gossip protocol would have
replicated all of the critical pieces so the swarm should survive the loss of
the seed and continue to completetion. Conversely, swarms relying on local 
knowledge only would almost certainly die in this situation. Finally, the
relatively small number of test swarms that were simulated in our current
results detracts somewhat from their realism. In the future, we would like
to run some addition test swarms to futher strengthen the support for our
findings.

We were originally interested in implementing gossip in the Vuze BitTorrent client
and running a simulation in PlanetLab. We decided against that because our simulator
gives easy access to global knowledge to evaluate the upper bound of performance improvement
(which turned out to be very important to our results). Now that we have found a practical
application of gossip in saving unhealthy swarms, we would like to go back and implement
the protocol in a real BitTorrent client and see how this does with real swarms.


Bram Cohen's original BitTorrent paper said that global knowledge is too expensive to acquire,
but we have shown that we can acquire a reasonable approximation of global knowledge
for certain types of knowledge in BitTorrent. Therefore, we would like to investigate
further appliations of improved global knowledge in BitTorrent. We are particularly interested in 
finding misbehaving nodes (e.g. identifying BitThief clients by gossiping about optimistic unchokes)
or other ways to improve performance (e.g. gossiping about the download rates achieved by nodes).
One complicating factor is that many possible uses of gossip also create interesting incentives
that need to be analyzed.

